Networks Matter in International Relations:\\A Latent Network Approach to Dyadic Analysis

### 2017-10-24 paper/1_intro.tex
edited manuscript

\citet{aronow:etal:2015} estimate that over the period from 2010 to 2015, over sixty articles utilizing dyadic data were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization}. Most of these studies use a generalized linear model (GLM) to estimate regression coefficients.  However, extant approaches to studying dyadic data increase the chance of faulty inferences by treating data as independent and identically distributed (iid) when observations may be highly dependent. Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters in the link function, so as to reflect the potential clustering of cases. This may work in limited situations, but is not generally effective because these palliatives do not address the fundamental data generating process that remains a threat to inference because of the interdependence of observations or measurements. Namely, it is not just the diagonals of the variance-covariance matrix this affects. % \citep{beck:2012, king:roberts:2014}

In this article, we discuss a Bayesian approach, the Additive and Multiplicative Effects (AME) model, for directly modeling relational data to reflect the data generating process that yields interdependencies in these types of data structures \citep{hoff:2008,minhas:etal:2016:arxiv}. We focus on three types of interdependencies that can complicate dyadic analyses. First, dependencies may arise within a set of dyads as there may be a particular actor that is more likely to send or receive events such as conflict. Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad; specifically, if a rebel group initiates a conflict with a government, the government will likely reciprocate that conflictual behavior. We capture these two dependencies, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model. The multiplicative effects capture dependencies that result from groups of actors clustering together or organizing into communities due to \textit{meso-scopic} features of networks, such as homophily and stochastic equivalence. These type of meso-scopic features often arise in relational data because actors possess some latent set of shared attributes that affect their probability of interacting with one another.

We begin with a brief review of these dependencies and the AME model. Next, we conduct a simulation study to show how this AME approach can recover unbiased and well-calibrated regression coefficients in the presence of network dependencies. Then, we apply this approach to five prominent studies in the international relations (IR) literature and compare results from the current state-of-the-art approach (a GLM with robust standard errors) to those obtained with the AME framework. The latent factor approach (AME) is able to better capture first, second, and third-order interdependencies than the standard approach. It also produces results that are, at times, at odds with those found in these studies in particular, and the broader literatures from which they are drawn. As such, this approach offers substantive insights that are often occluded by ignoring the interdependencies in most relational data that the field of IR is concerned with. Finally, we show that for each replication our network based approach provides substantively more accurate out-of-sample predictions than the models used in the original studies. Thus, the AME approach is one that can be used by scholars in the field to not only generate substantive insights, but it also enables us to better model the data generating process behind events of interest.  Most importantly, it facilitates the concentration on the relations aspect of the field of international relations.

>>>>

\citet{aronow:etal:2015} estimate that over the period from 2010 to 2015, over sixty articles utilizing dyadic data were published in the \textit{American Political Science Review}, \textit{American Journal of Political Science}, and \textit{International Organization}. Most of these studies use a generalized linear model (GLM) to estimate regression coefficients.  However, this approach to studying dyadic data increase the chance of faulty inferences by treating data as independent and identically distributed (iid) when observations may be highly dependent. Most standard approaches assume that the problems raised by having non-iid relational data can be addressed by recalculating the standard errors of estimated parameters in the link function, so as to reflect the potential clustering of cases. This may work in limited situations, but is not generally effective because these palliatives do not address the fundamental data generating process that remains a threat to inference because of the interdependence of observations or measurements. Namely, it is not just the diagonals of the variance-covariance matrix this affects. % \citep{beck:2012, king:roberts:2014}

In this article, we discuss a Bayesian approach, the Additive and Multiplicative Effects (AME) model, for directly modeling relational data to reflect the data generating process that yields interdependencies in these types of data structures \citep{hoff:2008,minhas:etal:2016:arxiv}. We focus on three types of interdependencies that can complicate dyadic analyses. First, dependencies may arise within a set of dyads as there may be a particular actor that is more likely to send or receive events such as conflict. Additionally, if the event of interest has a clear sender and receiver, we are likely to observe dependencies within a dyad; for example, if a rebel group initiates a conflict with a government, the government will likely reciprocate that behavior. We capture these two dependencies, often referred to as first- and second-order dependencies, respectively, within the additive effects portion of the model. The multiplicative effects capture dependencies that result from groups of actors clustering together or organizing into communities due to \textit{meso-scopic} features of networks, such as homophily and stochastic equivalence. These type of meso-scopic features often arise in relational data because actors possess some latent set of shared attributes that affect their probability of interacting with one another.

We begin with a brief review of these dependencies and the AME model. Next, we conduct a simulation study to show how this AME approach can recover unbiased and well-calibrated regression coefficients in the presence of network dependencies. Then, we apply this approach to five prominent studies in the international relations (IR) literature and compare results from the current state-of-the-art approach (a GLM with robust standard errors) to those obtained with the AME framework. The latent factor approach (AME) is better able to capture first, second, and third-order interdependencies than the standard approach. It also produces results that are, at times, at odds with those found in these studies in particular, and the broader literatures from which they are drawn. As such, this approach offers substantive insights that are often occluded by ignoring the interdependencies in most relational data that the field of IR is concerned with. Finally, we show that for each replication our network based approach provides substantively more accurate out-of-sample predictions than the models used in the original studies. Thus, the AME approach is one that can be used by scholars in the field to not only generate substantive insights, but it also enables us to better model the data generating process behind events of interest.  Most importantly, it facilitates concentration on the relations aspect of the field of international relations.

### 2017-10-24 paper/2_model.tex
edited manuscript

In working with relational data, scholars in the field begin by structuring it as a set of dyadic observations stacked on top of one another. This makes sense if each observation is independent of the others. Thus, for example, a conflict sent from the United States to Japan, is assumed to be independent of any action that Japan may send to the United States. Additionally, every action sent by Japan to others in the system is considered independent even though each of those interactions involves a common sender, i.e, Japan. As a result, the assumption that most begin with is that each dyadic interaction is taking place in isolation of the others. We know this assumption to be false in theory and in practice because relational data comes with an explicit structure that in general leads to particular types of dependencies. The importance of accounting for the underlying structure of our data has been a lesson well understood, at least, when it comes to time-series cross-sectional data (TSCS) within political science \citep{beck:katz:1995,beck:etal:1998}. As a result, it is now standard practice to take explicit steps to account for the complex data structures that emerge in TSCS applications and the unobserved heterogeneity that they cause. To uncover the underlying structure that relational data comes with, it is helpful to shift towards restructuring dyadic data in the form of a matrix --- often referred to as an adjacency matrix --- as shown in Figure~\ref{fig:adjMatDeps}. Rows designate the senders of an event and columns the receivers. The cross-sections in this matrix represent the actions that were sent by an actor in the row to those designated in the columns. Thus $y_{ij}$ designates an action $y$, such as a conflictual event or trade flows, that is sent from actor $i$ to actor $j$.

@@

\begin{align}
	\begin{aligned}
		y_{ij} \;=\; f(\theta_{ij}) &\text{, where } \\
		\theta_{ij} \;=\;& \bm\beta_{d}^{\top} \mathbf{X}_{ij} + \bm\beta_{s}^{\top} \mathbf{X}_{i} + \bm\beta_{r}^{\top} \mathbf{X}_{j} + \text{\qquad(Exogenous parameters)} \\
		& a_{i} + b_{j} + \epsilon_{ij} + \text{\qquad\qquad\qquad\quad(SRRM parameters)} \\
		& \textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}  \text{\qquad\qquad\qquad\qquad\qquad\;(LFM parameters)} \\ 
	\label{eqn:ame}
	\end{aligned}
\end{align}

where $y_{ij,t}$ captures the interaction between actor $i$ (the sender) and $j$ (the receiver). We use a Bayesian probit framework, in which we model a latent variable, $\theta_{ij}$, using first a set of exogenous dyadic ($\bm\beta_{d}^{\top} \mathbf{X}_{ij}$), sender ($\bm\beta_{s}^{\top} \mathbf{X}_{i}$), and receiver covariates ($\bm\beta_{r}^{\top} \mathbf{X}_{j}$). Next, to account for the dependencies that emerge in dyadic data and that may complicate inference on the parameter associated with exogenous covariates, we add parameters from the SRRM and LFM. $a_{i}$ and $b_{j}$ in Equation~\ref{eqn:ame} represent sender and receiver random effects incorporated from the SRRM framework:

>>>>

In working with relational data, scholars in the field begin by structuring it as a set of dyadic observations stacked on top of one another. This makes sense if each observation is independent of the others. For example, a conflict sent from the United States to Japan, is assumed to be independent of any action that Japan may send to the United States. Additionally, every action sent by Japan to others in the system is considered independent even though each of those interactions involves a common sender, i.e, Japan. While the assumption that most begin with is that each dyadic interaction is taking place in isolation of the others, we know this assumption to be false in theory and in practice. Relational data comes with an explicit structure that in general leads to particular types of dependencies. The importance of accounting for the underlying structure of our data has been a lesson well understood, at least, when it comes to time-series cross-sectional data (TSCS) within political science \citep{beck:katz:1995,beck:etal:1998}. As a result, it is now standard practice to take explicit steps to account for the complex data structures that emerge in TSCS applications and the unobserved heterogeneity that they cause. To uncover the underlying structure that relational data comes with, it is helpful to shift towards restructuring dyadic data in the form of a matrix --- often referred to as an adjacency matrix --- as shown in Figure~\ref{fig:adjMatDeps}. Rows designate the senders of an event and columns the receivers. The cross-sections in this matrix represent the actions that were sent by an actor in the row to those designated in the columns. Thus $y_{ij}$ designates an action $y$, such as a conflictual event or trade flow, that is sent from actor $i$ to actor $j$.

@@

\begin{align}
	\begin{aligned}
		y_{ij} \;=\; f(\theta_{ij}) &\text{, where } \\
		\theta_{ij} \;=\;& \bm\beta_{d}^{\top} \mathbf{X}_{ij} + \bm\beta_{s}^{\top} \mathbf{X}_{i} + \bm\beta_{r}^{\top} \mathbf{X}_{j} + \text{\qquad(Exogenous parameters)} \\
		& a_{i} + b_{j} + \epsilon_{ij} + \text{\qquad\qquad\qquad\quad(SRRM parameters)} \\
		& \textbf{u}_{i}^{\top} \textbf{D} \textbf{v}_{j}  \text{\qquad\qquad\qquad\qquad\qquad\;(LFM parameters)} \\ 
	\label{eqn:ame}
	\end{aligned}
\end{align}
where $y_{ij,t}$ captures the interaction between actor $i$ (the sender) and $j$ (the receiver). We use a Bayesian probit framework, in which we model a latent variable, $\theta_{ij}$, using first a set of exogenous dyadic ($\bm\beta_{d}^{\top} \mathbf{X}_{ij}$), sender ($\bm\beta_{s}^{\top} \mathbf{X}_{i}$), and receiver covariates ($\bm\beta_{r}^{\top} \mathbf{X}_{j}$). Next, to account for the dependencies that emerge in dyadic data and that may complicate inference on the parameter associated with exogenous covariates, we add parameters from the SRRM and LFM. $a_{i}$ and $b_{j}$ in Equation~\ref{eqn:ame} represent sender and receiver random effects incorporated from the SRRM framework:

### 2017-10-24 paper/3_sim.tex
edited manuscript

\begin{align}
	y_{i,j} \sim  \mu + \beta x_{i,j} + \gamma w_{i,j} + \epsilon_{i,j}
	\label{eqn:sim}
\end{align}

where $Y= \{y_{i,j}\}\in \mathbb R^{n\times n}$ is an observed sociomatrix, $X = \{x_{i,j} \} \in \mathbb R^{n \times n}$ is a matrix of observed dyad-specific characteristics, and $W = \{ w_{i,j}\} \in \mathbb R^{n \times n}$ is a matrix of unobserved dyad-specific characteristics. $Y$ can be thought of as a dyadic dependent variable, $X$ and $W$ are both dyadic covariates that are a part of the data-generating process for $Y$, but $W$ is a variable that we do not observe. We compare inference for $\mu$ and $\beta$ using three models --- the latter would be of primarily theoretical concern for applied scholars:

@@

Next, we estimate the 95\% confidence interval for the three models in each of the simulations and estimate the proportion of times that the true value actually fell in those intervals. The results are summarized in Figure~\ref{fig:ameCalib}, and here again we see that the AME approach performs as well as the oracle while the naive approach performs poorly.

>>>>

\begin{align}
	y_{i,j} \sim  \mu + \beta x_{i,j} + \gamma w_{i,j} + \epsilon_{i,j}
	\label{eqn:sim}
\end{align}
where $Y= \{y_{i,j}\}\in \mathbb R^{n\times n}$ is an observed sociomatrix, $X = \{x_{i,j} \} \in \mathbb R^{n \times n}$ is a matrix of observed dyad-specific characteristics, and $W = \{ w_{i,j}\} \in \mathbb R^{n \times n}$ is a matrix of unobserved dyad-specific characteristics. $Y$ can be thought of as a dyadic dependent variable, $X$ and $W$ are both dyadic covariates that are a part of the data-generating process for $Y$, but $W$ is a variable that we do not observe. We compare inference for $\mu$ and $\beta$ -- the latter parameter would be of primarily theoretical concern for applied scholars -- using three models :

@@

Next, we estimate the 95\% confidence interval for the three models in each of the simulations and estimate the proportion of times that the true value actually fell in those intervals. The results are summarized in Figure~\ref{fig:ameCalib}, and again we see that the AME approach performs as well as the oracle while the naive approach performs poorly.

### 2017-10-24 paper/4_app.tex
edited manuscript

% SM: maybe too on the nose (Whether putative control variables, for example, are precisely estimated is less important.) 
Given current practices, most scholars have a single variable in a complicated empirical model that they look at for evaluating the validity of their empirically estimated models. Generally, interpretations focus on a small set of independent variables.  In the study by Reiter and Stam (2003) the goal is to determine in mixed dyads --- consisting of a democratic country and country ruled by a personalist dictatorship --- whether it is the democratic country or the personalist country that is associated with a greater likelihood of the dyad being in a dispute.  In a regression including fourteen covariates, they focus on whether the variables that capture whether the initiator is a democratic country are statistically significant.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our re-examination of Reiter and Stam, we focus on whether the model replicates when re-estimated within the AME framework. Each of the other four studies also has a crucial finding that we hone in on to further draw into focus the potential differences brought about by utilizing an AME estimation procedure.  In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{does not replicate} indicates only that the sign and/or significance of the putatively crucial finding in the original study is not found to hold in the AME estimation.

\begin{table}[ht]
\centering
\caption{Here we provide a brief summary of the key variable in each of the five replications and a note about whether or not the finding is replicated when using our network based approach. Cases in which the substantive finding is not replicated are highlighted in bold.}
	\begin{tabular}{l p{7cm} l} \toprule
		\multirow{2}{*}{Study} & \multirow{2}{*}{Central Finding} &  Replicates \\ 
			& &  in a Network Model? \\ \toprule
		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Replicates} \\ \midrule
		McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Does Not Replicate}\\ \midrule
		Rose (2004) & WTO Membership Does not Effect Trade & {Replicates}\\ \midrule
		Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & {Does Not Replicate} \\\midrule
		Gibler (2017) & Power Parity at Time of Entry to International System Inceases Conflict & {Does Not Replicate} \\ \bottomrule
	\end{tabular}
	\label{tab:modelFindingSumm}
\end{table}

In addition, we also examine the predictions made with each approach using an out-of-sample cross validation strategy. By accounting for exogenous and network dependent patterns that give rise to conflict systems we are able to better account for the data generating process underlying relational data structures. To show that this is the case, we examine whether our approach achieves better predictive performance in an out-of-sample context than traditional dyadic models. To evaluate our model, we randomly divide the $\binom n 2 \times T$ data values into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned. Then for each $s \in \{1,\ldots,k\}$, we:

@@

\begin{table}[ht]
\centering
	\begin{tabular}{l|l cc}
	~ & ~ & GLM & AME \\
	\toprule
\multirow{2}{*}{Reiter \& Stam (2003)} & Area Under ROC Curve, AUC-ROC & 0.92 & {0.96} \\
		~ & Area Under PR Curve, AUC-PR & 0.08 &  {0.15} \\		\midrule
\multirow{2}{*}{McDonald (2004)} & AUC-ROC & 0.92 &  {0.99} \\
		~ & AUC-PR & 0.13 &  {0.28} \\		\midrule
\multirow{2}{*}{Rose (2004)} & RMSE & 3.23 &  {1.99} \\
		~ & RMDSE & 2.01 &  {1.06} \\	\midrule
\multirow{2}{*}{Weeks (2012)} & AUC-ROC & 0.64 &  {0.97} \\
		~ & AUC-PR & 0.00 &  {0.15} \\		\midrule
\multirow{2}{*}{Gibler (2017)} & AUC-ROC & 0.52 &  {0.91} \\
		~ & AUC-PR & 0.00 &  {0.08} \\			\bottomrule
	\end{tabular}
	\caption{Here we provide a summary of the out-of-sample performance based on our cross-validation strategy for each of the five replications when using the standard dyadic approach and our network based approach. Four of the five studies involved a binary dependent variable, so for those measures we provide area under the curve (AUC) statistics. The Rose study involved a Gaussian dependent variable and for that we use the root mean squared error (RMSE) and root median squared error (RMDSE). Cases in which our network based approach outperformed the standard approach are highlighted in bold.}
	\label{tab:modelPerfSumm}
\end{table}
\FloatBarrier

@@

In 2004, Andrew Rose published a study in the \textit{American Economic Review} \nocite{rose:2004} that proved to be quite controversial in terms of macroeconomic trade theory and in terms of trade policy in a variety of nations. It also provoked a number of responses in the international political economy literature \cite{tomz:etal:2007,ward:etal:2013}.  Rose's basic argument is that despite longstanding arguments made by trade theorists and the World Trade Organization that WTO membership fosters greater cooperation and thereby more trade among its members, the empirics do not bear out such claims. He uses a standard gravity model with dyadic data on bilateral merchandise trade (not services) for $175$ countries over a period of five decades. Estimating this model using OLS within many differing contexts, his conclusion was that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders \ldots (2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO --- as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.

When we compare the results of Rose's original OLS model to the AME model accounting for network dependencies, the results are generally similar. The main result of the model --- the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables --- persists in the AME model. However, while in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model. The random effects shown in Figure~\ref{fig:roser} reveal the cause of much of this divergence. Here, the states with the most positive random effects are also states with high GDP, though not necessarily high GDP/capita.\footnote{Note: Qatar exhibits strongly negative random effects.} Thus, the effect of GDP in the original model was, in part, an artifact of first-order dependencies. Most of the other results of the model are constant across each model, though some geographic features, such as islands and landlocked states, have a more clear effect on trade once we account for these network dependencies. When we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out-of-sample --- $3.23$ for the OLS model and $1.77$ for the AME model. While the basic model replicates with the basic result, there are substantial and substantive improvements gleaned from the AME results.

@@

\citet{weeks:2012} examines the influence of domestic institutions on the initiation of military conflicts by autocratic leaders.  She argues that in some circumstances autocrats are held accountable for their foreign policy decisions. She adds the nuance that autocratic audiences are not homogeneous. When the autocratic regime is nonmilitary, the domestic audience do not favor military actions, but in military autocracies this is not the case. Further she argues that in personalistic regimes without a military or civilian domestic audience, the leaders tend to be more likely to employ military force in their foreign policy.  To study this question, she uses a dyadic design in which the dependent variable is ``whether country A in a directed dyad initiated military conflict against country B during year t'' (page 337).  These data come from the Militarized Interstate Disputes database \citep{maoz:2005}.  One major innovation in her study resides in the nuanced way in which she conceptualized and coded regime type into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen. She also includes a variety of putative control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.''  She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with  fixed, unit effects. The analysis is done for dyads, but is considered to be from the perspective of the actor that initiated the dispute. Her time frame is from 1946-1999.  The basic finding is that a) juntas, bosses, and strongmen are more likely to initiate conflict than machines (and maybe democracies) and that b) machines are no more belligerent than democracies.  These insights are mainly determined by looking at coefficients and their putative significance in Tables 1 and 2 (pages 339--340). She makes the argument that ignoring important nuances between different types of autocracies hinders our understanding of the initiation of military conflict by autocracies.

@@

The re-estimation of \citet{weeks:2012} likely has the sharpest divergence between the General Linear Model results and those of the AME Model. In Weeks's initial models,  she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses and Strong-men are more conflict-prone, in line with her theoretical expectations. When we look at the posterior distribution of these coefficients in the AME results, we find that none of these values are distinguishable from zero. We similarly find less pronounced effects for military capabilities. One explanation for this divergence is the AME model's ability to account for third order effects. Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence --- in the top right corner we see states like the US, the UK, and Israel who often send conflict to similar targets. In particular, we observe a cluster of ``rogue states,'' who are receivers of conflict, in the top right (Iran, Iraq, Syria, Libya, North Korea). In the bottom corner we see a cluster of authoritarian senders including Iraq, Russia, Syria, North Korea and China. In general, these clusters have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category), and similar conflict behavior. In the GLM, which ignores these third order dependencies, many of these results might have been attributed to regime type. The AME model, on the other hand, shows that this can be specified in terms of the interdependencies captured by the multiplicative effects. In terms of out-of-sample performance, shown in Figure~\ref{fig:weeksauc}, the AME model performs markedly better than the GLM out of sample, lending credence to the possibility that these third-order dependencies cause spurious effects for regime type.  In terms of performance, the original Weeks' model is slightly better than a coin-toss overall, and fails to indicate more than a couple of events (incorrectly) in almost one million observations.

@@

First and foremost, many findings which emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.\footnote{This should not be news --- since the finding has long been in the theoretical literature, but given the state of current literature in international relations is still pertinent.} Not only are coefficients biased in the OLS and logit approaches to the analysis of dyadic data, but they are often imprecisely measured, with inflated standard errors.  This means that significant testing (for better or worse) is compromised when network effects are ignored.

@@

Fourth, it is clear that the AME approach dominates the OLS and Logit-based approaches in terms of performance. Not only it is better at correctly identifying cases in which the dependent variable takes a value of $0$ (via the ROC curves and associated statistics), but it also dominates at correctly identifying occurrences of the dependent variable in the data (seen via the PR curves and associated statistics).  In the case of studies with continuous dependent variables, the AME approach has average error statistics that are about one-half that found in the OLS model. It is rare for studies in this area to provide performance statistics, but at the same time that at least one of the studies is unable to identify a single case in spite of having almost a million observations.

>>>>

% SM: maybe too on the nose (Whether putative control variables, for example, are precisely estimated is less important.) 
Given current practices, most scholars have a single variable in a complicated empirical model that they look at for evaluating the validity of their empirically estimated models. Generally, interpretations focus on a small set of independent variables.  For example, in the study by Reiter and Stam (2003) the goal is to determine whether in mixed dyads --- consisting of a democratic country and country ruled by a personalist dictatorship --- it is the democratic country or the personalist country that is associated with a greater likelihood of the dyad being in a dispute.  In a regression including fourteen covariates, they focus on whether the variables that capture whether the initiator is a democratic country are statistically significant.\footnote{In particular, the variables \texttt{Pers/Democ} and \texttt{Democ/Pers} in Model 3, in Table 1, page 335.} In our re-examination of Reiter and Stam, we focus on whether the model replicates when re-estimated within the AME framework. Each of the other four studies also has a crucial finding that we hone in on to further draw into focus the potential differences brought about by utilizing an AME estimation procedure.  In Table~\ref{tab:modelFindingSumm}, we present the overall results; the term \textit{does not replicate} indicates only that the sign and/or significance of the putatively crucial finding in the original study is not found to hold in the AME estimation.

\begin{table}[ht]
\centering
\caption{Here we provide a brief summary of the key variable in each of the five replications and a note about whether or not the finding is replicated when using our network based approach. Cases in which the substantive finding is not replicated are highlighted in bold.}
	\begin{tabular}{l p{7cm} l} \toprule
		\multirow{2}{*}{Study} & \multirow{2}{*}{Central Finding} &  Replicates \\ 
			& &  in a Network Model? \\ \toprule
		Reiter \& Stam (2003) & Personalist Regimes Attack Democracies, Not Vice Versa & {Partially Replicates} \\ \midrule
		\bf{McDonald (2004) & Lower Trade Barriers and Higher Trade Lead to Peace & {Does Not Replicate}}\\ \midrule
		Rose (2004) & WTO Membership Does not Effect Trade & {Replicates}\\ \midrule
		\bf{Weeks (2012) & Bosses, Juntas and Strongmen are more Aggressive, Machines are Not & {Does Not Replicate}} \\\midrule
		\bf{Gibler (2017) & Power Parity at Time of Entry to International System Inceases Conflict & {Does Not Replicate} }\\ \bottomrule
	\end{tabular}
	\label{tab:modelFindingSumm}
\end{table}

In addition, we also examine the predictions made with each approach using an out-of-sample cross validation strategy. By accounting for exogenous and network dependent patterns that give rise to conflict systems we are able to better account for the data generating process underlying relational data structures. To show that this is the case, we examine whether our approach achieves better predictive performance in an out-of-sample context than traditional dyadic models. To evaluate our model, we randomly divide the data  into $k=30$ sets, letting $s_{ij,t}$ be the set to which pair $ij,t$ is assigned.\textcolor{red}{This is inaccurate for directed dyads. I think. MG} Then for each $s \in \{1,\ldots,k\}$, we:

@@

\begin{table}[ht]
\centering
	\begin{tabular}{l|l cc}
	~ & ~ & GLM & AME \\
	\toprule
\multirow{2}{*}{Reiter \& Stam (2003)} & Area Under ROC Curve, AUC-ROC & 0.92 & {0.96} \\
		~ & Area Under PR Curve, AUC-PR & 0.08 &  {0.15} \\		\midrule
\multirow{2}{*}{McDonald (2004)} & AUC-ROC & 0.92 &  {0.99} \\
		~ & AUC-PR & 0.13 &  {0.28} \\		\midrule
\multirow{2}{*}{Rose (2004)} & RMSE & 3.23 &  {1.99} \\
		~ & RMDSE & 2.01 &  {1.06} \\	\midrule
\multirow{2}{*}{Weeks (2012)} & AUC-ROC & 0.64 &  {0.97} \\
		~ & AUC-PR & 0.00 &  {0.15} \\		\midrule
\multirow{2}{*}{Gibler (2017)} & AUC-ROC & 0.52 &  {0.91} \\
		~ & AUC-PR & 0.00 &  {0.08} \\			\bottomrule
	\end{tabular}
	\caption{Here we provide a summary of the out-of-sample performance based on our cross-validation strategy for each of the five replications when using the standard dyadic approach and our network based approach. Four of the five studies involved a binary dependent variable, so for those measures we provide area under the curve (AUC) statistics. The Rose study involved a Gaussian dependent variable and for that we use the root mean squared error (RMSE) and root median squared error (RMDSE). %Cases in which our network based approach outperformed the standard approach are highlighted in bold.
	}
	\label{tab:modelPerfSumm}
\end{table}
\FloatBarrier

@@

In 2004, Andrew Rose published a study in the \textit{American Economic Review} \nocite{rose:2004} that proved to be quite controversial in terms of macroeconomic trade theory and in terms of trade policy in a variety of nations. It also provoked a number of responses in the international political economy literature \cite{tomz:etal:2007,ward:etal:2013}.  Rose's basic argument is that despite longstanding arguments, made by trade theorists and the World Trade Organization, that WTO membership fosters greater cooperation and thereby more trade among its members, the empirics do not bear out such claims. He uses a standard gravity model with dyadic data on bilateral merchandise trade (not services) for $175$ countries over a period of five decades. Estimating this model using OLS within many differing contexts, his conclusion was that: ``An extensive search reveals little evidence that countries joining or belonging to the GATT/WTO have different trade patters from outsiders \ldots (2004, page 98, abstract).''  The data for this study have been widely used in replications by many searching for the missing effects of the WTO --- as well as preferential trade agreements, bilateral investment theories, and other aspects of modern trade theory.

When we compare the results of Rose's original OLS model to the AME model accounting for network dependencies, the results are generally similar. The main result of the model --- the null effect of membership in the WTO, as represented by the ``One-In" and ``Both-In" variables --- persists in the AME model. However, while in the original model there was a clear positive relationship between Real GDP and Trade, most of this effect vanishes in the AME model. The random effects shown in Figure~\ref{fig:roser} reveal the cause of much of this divergence. Here, the states with the most positive random effects are also states with high GDP, though not necessarily high GDP/capita.\footnote{Note: Qatar exhibits strongly negative random effects.} Thus, the effect of GDP in the original model was, in part, an artifact of first-order dependencies. Most of the other results of the model are constant across each model, though some geographic features, such as islands and landlocked states, have a more clear effect on trade once we account for these network dependencies. When we account for network interdependencies, we observe a markedly lower Root Mean Squared Error out-of-sample --- $3.23$ for the OLS model and $1.77$ for the AME model. While the AME model replicate the main result, there are substantial and substantive improvements gleaned from the AME results.

@@

\citet{weeks:2012} examines the influence of domestic institutions on the initiation of military conflicts by autocratic leaders.  She argues that in some circumstances autocrats are held accountable for their foreign policy decisions. She adds the nuance that autocratic audiences are not homogeneous. When the autocratic regime is nonmilitary, the domestic audience do not favor military actions, but in military autocracies this is not the case. Further she argues that in personalistic regimes without a military or civilian domestic audience, the leaders tend to be more likely to employ military force in their foreign policy.  To study this question, she uses a dyadic design in which the dependent variable is ``whether country A in a directed dyad initiated military conflict against country B during year t'' (page 337).  These data come from the Militarized Interstate Disputes database \citep{maoz:2005}.  One major innovation in her study resides in the nuanced way in which she conceptualized and coded regime type into four types: a) Machine, b) Junta, c) Boss, and d) Strongmen. She also includes a variety of putative control variables focusing on capabilities for both sides of the dyad, alliances, geography, trade dependence, regime instability, and the regime type of ``side B.''  She uses a logistic regression, but follows \citet{beck:etal:1998} and includes splines to capture temporal covariation in the dependent variable along with  fixed, unit effects. The analysis is done for dyads, but is considered to be from the perspective of the actor that initiated the dispute. %Her time frame is from 1946-1999.  
The basic finding is that a) juntas, bosses, and strongmen are more likely to initiate conflict than machines (and maybe democracies) and that b) machines are no more belligerent than democracies.  These insights are mainly determined by looking at coefficients and their putative significance in Tables 1 and 2 (pages 339--340). She makes the argument that ignoring important nuances between different types of autocracies hinders our understanding of the initiation of military conflict by autocracies.

@@

The re-estimation of \citet{weeks:2012} likely has the sharpest divergence between the General Linear Model results and those of the AME Model. In Weeks's initial models,  she finds that machines are less prone to initiate conflict than the reference category, whereas Juntas, Bosses and Strong-men are more conflict-prone, in line with her theoretical expectations. When we look at the posterior distribution of these coefficients in the AME results, we find that none of these values are distinguishable from zero. Similarly, we find less pronounced effects for military capabilities. One explanation for this divergence is the AME model's ability to account for third order effects. Inspection of the multiplicative effects in Figure~\ref{fig:weekscirc} reveals a number of clusters of states which exhibit structural equivalence --- in the top right corner we see states like the US, the UK, and Israel who often send conflict to similar targets. We also observe a cluster of ``rogue states,'' who are receivers of conflict, in the top right (Iran, Iraq, Syria, Libya, North Korea). In the bottom corner we see a cluster of authoritarian senders including Iraq, Russia, Syria, North Korea and China. In general, these clusters have similar governmental types (Iraq, Syria, Libya, and North Korea all fell under the ``boss" category), and similar conflict behavior. In the GLM, which ignores these third order dependencies, many of these results might have been attributed to regime type. The AME model, on the other hand, shows that this can be specified in terms of the interdependencies captured by the multiplicative effects. In terms of out-of-sample performance, shown in Figure~\ref{fig:weeksauc}, the AME model performs markedly better than the GLM out of sample, lending credence to the possibility that these third-order dependencies cause spurious effects for regime type.  In terms of performance, the original Weeks' model does slightly better than a coin-toss overall, and fails to predict more than a couple of events (incorrectly) in almost one million observations.

@@

First and foremost, many findings which emerge from models that do not take interdependencies into account lose their statistical significance when network effects are estimated via AME.\footnote{This should not be news --- since the finding has long been in the theoretical literature, but given the state of current literature in international relations is still pertinent.} Not only are coefficients biased in the GLM approaches to the analysis of dyadic data, but they are often imprecisely measured, with inflated standard errors.  This means that significant testing (for better or worse) is compromised when network effects are ignored.

@@

Fourth, it is clear that the AME approach dominates the GLM approaches in terms of performance. Not only it is better at correctly identifying cases in which the dependent variable takes a value of $0$ (via the ROC curves and associated statistics), but it also dominates at correctly identifying occurrences of the dependent variable in the data (seen via the PR curves and associated statistics).  In the case of studies with continuous dependent variables, the AME approach has average error statistics that are about one-half that found in the OLS model. It is rare for studies in this area to provide performance statistics, but at the same time that at least one of the studies is unable to identify a single case in spite of having almost a million observations.
